\vspace{-.5cm} Lights: a generalized joint model for high-dimensional multivariate longitudinal data and censored durations \vspace{.5cm}

### 2021-08-19 paper.tex
get version 1 of the lemma 1 proof, to be refactored

\newcommand{\dd}{\mathrm{d}}
\newcommand{\ind}[1]{\mathds{1}_{#1}}
\newcommand{\norm}[1]{\|#1\|}
\newcommand{\cM}{\mathcal M}
\newcommand{\cD}{\mathcal D}
\newcommand{\cC}{\mathcal C}
\newcommand{\cG}{\mathcal G}
\newcommand{\cN}{\mathcal N}
\newcommand{\cA}{\mathcal A}
\newcommand{\sA}{\mathscr A}
\newcommand{\cQ}{\mathcal Q}
\newcommand{\cU}{\mathcal U}
\newcommand{\cR}{\mathcal R}
\newcommand{\cH}{\mathcal H}
\newcommand{\cB}{\mathcal B}
\newcommand{\cS}{\mathcal S}
\newcommand{\R}{\mathds R}
\newcommand{\N}{\mathds N}
\newcommand{\E}{\mathds E}
\renewcommand{\P}{\mathds P}
\newcommand{\bSigma}{\textbf{$\Sigma$}}

@@

\texttt{define $\Sigma_i$}

@@

@@

Let us introduce the following useful notations
\begin{equation}
  \label{eq:g-tilde}
  \left\{
    \begin{array}{ll}
    \tilde g^1&: b_i \mapsto b_i,\\
    \tilde g^2&: b_i \mapsto b_ib_i^\top, \\
    \tilde g^3_{s}&: b_i \mapsto \Psi_i(s), \\
    \tilde g^4_{s, \gamma_k}&: b_i \mapsto \exp \big\{\gamma_k^\top g^3_{s}(b_i) \big\}, \\
    \tilde g^5_{s, \gamma_k}&: b_i \mapsto g^3_{s}(b_i)^\top g^4_{s}(b_i).
    \end{array}
  \right. 
\end{equation}

>>>>

\newcommand{\dd}{\mathrm{d}}
\newcommand{\ind}[1]{\mathds{1}_{#1}}
\newcommand{\norm}[1]{\|#1\|}
\newcommand{\cM}{\mathcal M}
\newcommand{\cD}{\mathcal D}
\newcommand{\cC}{\mathcal C}
\newcommand{\cG}{\mathcal G}
\newcommand{\cN}{\mathcal N}
\newcommand{\cA}{\mathcal A}
\newcommand{\sA}{\mathscr A}
\newcommand{\cQ}{\mathcal Q}
\newcommand{\cU}{\mathcal U}
\newcommand{\cR}{\mathcal R}
\newcommand{\cH}{\mathcal H}
\newcommand{\cB}{\mathcal B}
\newcommand{\cS}{\mathcal S}
\newcommand{\cI}{\mathcal I}
\newcommand{\R}{\mathds R}
\newcommand{\N}{\mathds N}
\newcommand{\E}{\mathds E}
\renewcommand{\P}{\mathds P}
\newcommand{\bSigma}{\textbf{$\Sigma$}}

@@

\texttt{TODO Van Tuan : define $\Sigma_i$}

@@

\texttt{TODO Simon : rewrite the proof properly}

The proximal operator for any function $f : R^p \to R$ is defined by
\begin{equation}
    \label{eq:prox-def}
    \text{prox}_{f}(u) = \argmin_{x \in \R^p} \{\frac{1}{2}\norm{x - u}_2^2 + f(x)\}.
\end{equation}
\begin{lemma} 
The following holds
\begin{equation*}
    \text{prox}_{\lambda_1 \norm{\cdot}_1 + \lambda_2 \sum_{l=1}^L  \norm{\cdot}_{2,l}}
    = \text{prox}_{\lambda_2 \sum_{l=1}^L \norm{\cdot}_{2,l}} \circ \text{prox}_{\lambda_1 \norm{\cdot}_1}
\end{equation*}
where $prox_{\lambda_1 \norm{\cdot}_1 + \lambda_2 \sum_{l=1}^L \norm{\cdot}_2}$ is the proximal operator for the sparse group lasso regulariser, $prox_{\lambda_2 \sum_{l=1}^L \norm{\cdot}_2}$ is the proximal operator for the group lasso regulariser and $prox_{\lambda_1 \norm{\cdot}_1}$ is the proximal operator for the lasso regulariser. $\lambda_1 > 0$, $\lambda_2 > 0$ are regularization parameters, $\norm{\cdot}_{2,l}$ denotes the Euclidean norm on features in l-th group.
\end{lemma}

\begin{proof}
Denote 
\begin{equation*}
    k(x) =  \frac{1}{2}\norm{ x - \text{prox}_{\lambda_1 \norm{\cdot}_1}(\beta)}_2^2 + \lambda_2 \sum_{l=1}^L \norm{x}_{2,l}
\end{equation*}
then we have the proximal operator (see~\eqref{eq:prox-def}) for the group lasso regulariser with $u = \text{prox}_{\lambda_1 \norm{\cdot}_1}(\beta)$ is given by
\begin{equation*}
    \text{prox}_{\lambda_2 \sum_{l=1}^L \norm{\cdot}}\big(\text{prox}_{\lambda_1 \norm{\cdot}_1}(\beta)\big) = \argmin_x k(x).
\end{equation*}
The sub-differential \texttt{[REF]} of $k(\cdot)$ at $x$ is given by
\begin{equation*}
    \partial k(x) = x - \text{prox}_{\lambda_1 \norm{\cdot}_1}(\beta) + \partial( \lambda_2 \sum_{l=1}^L \norm{x}_{2,l} )
\end{equation*}
It is clear that $k(\cdot)$ is convex function. Denote the unique minimizer of $k(\cdot)$ is $x^\star$. The sufficient and necessary condition for the optimality of $x^\star$ is 
\begin{equation}
    \label{eq:opt_gl}
    0 \in x^* - \text{prox}_{\lambda_1 \norm{\cdot}_1}(\beta) + \partial( \lambda_2 \sum_{l=1}^L \norm{x^\star}_{2,l} ).
\end{equation}
\texttt{[REF]}.
Denote $y = \text{prox}_{\lambda_1\norm{\cdot}_1}(\beta)$. From Lemma 1 in ~\citet{yuan2011efficient}, one has that $\text{SGN}(y) \subseteq \text{SGN}(x^*)$\\
where \[\text{SGN}(x) = 
\left\{
  \begin{array}{ll}
    \{1\} &\text{if } x > 0, \\ 
    $[-1, 1]$ &\text{if } x = 0, \\ 
    \{-1\} &\text{if } x < 0. 
  \end{array}
\right. \]
which is the sub-differential of the l1 norm. \texttt{[REF]} \\
\noindent Denote 
\begin{equation*}
    h(x) =  \frac{1}{2}\norm{x - \beta}_2^2 + \lambda_1 \norm{x}_1 + \lambda_2 \sum_{l=1}^L \norm{x}_{2,l}
\end{equation*}
then we have the  proximal  operator  for  the  sparse  group  lasso  regulariser is given by
\begin{equation*}
    \text{prox}_{\lambda_1 \norm{\cdot}_1 + \lambda_2 \sum_{l=1}^L \norm{\cdot}}(\beta) = \argmin_{x} \{ h(x)\}.
\end{equation*}
Denote the sub-differential of $h(\cdot)$ at $x$ is given by
\begin{align*}
    \partial(h(x)) &= x - \beta + \partial(\lambda_1 \norm{x}_1) + \partial( \lambda_2 \sum_{l=1}^L \norm{x}_{2, l} ) \\
    &= x - \beta + \lambda_1 \text{SGN}(x) + \partial( \lambda_2 \sum_{l=1}^L \norm{x}_{2, l} )
\end{align*}

\noindent We need to prove that $0 \in \partial\big(h(x^\star)\big)$. We can easily prove that 
\begin{equation*}
    y = \text{prox}_{\lambda_1 \norm{\cdot}_1}(\beta) 
    = \argmin_{u} \{\norm{u - \beta}_2^2 + \lambda_1 \norm{u}_1\} 
    = \beta - \lambda_1 \text{SGN}(y) \in \beta - \lambda_1 \text{SGN}(x^*).
\end{equation*}
Hence
\begin{equation}
    \label{eq:opt_gl_in_sgl}
    x^\star - \text{prox}_{\lambda_1 \norm{\cdot}_1}(\beta) + \partial( \lambda_2 \sum_{l=1}^L \norm{x^\star}_{2, l} ) \in x^\star- \beta + \lambda_1 \text{SGN}(x^\star) + \partial( \lambda_2 \sum_{l=1}^L \norm{x^\star}_{2, l} )
\end{equation}
It follows from ~\eqref{eq:opt_gl}- ~\eqref{eq:opt_gl_in_sgl} that $0 \in \partial(h(x^\star))$. Hence
\begin{equation*}
    \text{prox}_{\lambda_1 \norm{\cdot}_1 + \lambda_2 \sum_{l=1}^L \norm{\cdot}} = \text{prox}_{\lambda_2 \sum_{l=1}^L \norm{\cdot}} \circ \text{prox}_{\lambda_1 \norm{\cdot}_1}
\end{equation*}
\end{proof}

@@

Let us introduce the following useful notations
\begin{equation}
  \label{eq:g-tilde}
  \left\{
    \begin{array}{ll}
    \tilde g^1&: b_i \mapsto b_i,\\
    \tilde g^2&: b_i \mapsto b_ib_i^\top, \\
    \tilde g^3_{s}&: b_i \mapsto \Psi_i(s), \\
    \tilde g^4_{s, \gamma_k}&: b_i \mapsto \exp \big\{\gamma_k^\top g^3_{s}(b_i) \big\}, \\
    \tilde g^5_{s, \gamma_k}&: b_i \mapsto g^3_{s}(b_i)^\top g^4_{s}(b_i).
    \end{array}
  \right. 
\end{equation}
We also denote 
\[ \cI_1 = 
\begin{bmatrix}
  \ind{\{t_{1} = t_1^u\}}  & \cdots & \ind{\{t_{1} = t_J^u\}}\\
  \vdots &  \ddots & \vdots \\
  \ind{\{t_{n} = t_1^u\}} & \cdots & \ind{\{t_{n} = t_J^u\}}
\end{bmatrix}
\in \R^{n \times J}
\quad \text{and} \quad 
\cI_2 = 
\begin{bmatrix}
  \ind{\{t_{1} \leq t_1^u\}} & \cdots & \ind{\{t_{1} \leq t_J^u\}}\\
  \vdots &  \ddots & \vdots \\
  \ind{\{t_{n} \leq t_1^u\}} & \cdots & \ind{\{t_{n} \leq t_J^u\}}
\end{bmatrix}
\in \R^{n \times J} \]
some matrices to compare each event times with the vector of distinct event times. Those matrices are widely used in practice to replace indicators. For instance
\begin{align*}
f(t_i, \delta_i| b_i, G_i = k ; \theta) &= \Big( \sum_{j=1}^J \lambda_0(t_j) \tilde g_{i, t_j, \gamma_k^, \beta_k}^1 \ind{\{t_j = t_i\}} \Big)^{\delta_i} \exp \Big( - \sum_{j=1}^J \lambda_0(t_j) \tilde g_{i, t_j, \gamma_k, \beta_k}^1 \ind{\{t_j \leq t_i\}} \Big) \\
& = \Big( \text{sum}(\mathbf{\lambda_0} \tilde g_{\gamma_k, \beta_k}^1 \cI_1) \Big)^{\delta_i} \exp \Big( - \text{sum} (\mathbf{\lambda_0} \tilde g_{\gamma_k, \beta_k}^1\cI_2) \Big)
\end{align*}

\texttt{TODO Van Tuan : rewrite the two previous lines addind the $^{(w)}$ (else it's not true ! since it's only ok for the estimator) + the rigth $\tilde g$ + introduce undefined notations.}
